End-User Software Customization by Direct Manipulation of Tabular Data

### 2020-08-15 paper.tex
final tweaks

Here we describe the three types of table adapters we have built so far
in Wildcard. These do not represent an exhaustive set of all possible
table adapters---in Section~\ref{sec:visions} we discuss other types of
adapters that would fit well into the general paradigm.

>>>>

\hypertarget{abstract-interface}{%
\subsubsection{Abstract interface}\label{abstract-interface}}

We begin by describing the abstract interface fulfilled by a table
adapter.

\emph{Returning a table}: A table adapter exposes a table of data: an
ordered list of records. Each record carries a unique identifier and
associates named attributes with values. Tables have a typed schema, so
the same attributes are shared across all records. We currently support
strings, numeric values, booleans, and datetimes as types. The columns
also carry some additional metadata, such as whether or not they are
read-only or editable.

A table adapter can update the contents of a table at any time in
response to changes in the underlying state (e.g., a DOM scraping
adapter can update the table when the page body changes). When data
changes, the query view is reactively updated in response.

\emph{Handling edits}: The query engine can issue a request to a table
adapter to make an edit to a record. The meaning of making an edit can
vary depending on the adapter: in the local storage adapter, a new value
may be persisted into local storage; in the DOM scraping adapter, an
edit may result in changing the value of a form field.

In addition, the query engine also sends additional information about
the combined query view to each table adapter:

\emph{Sorting/filtering}: When the user sorts or filters the query view,
an ordered list of visible IDs is sent to each table adapter. The DOM
scraping adapter uses this information to change the list of rows shown
in the web page.

\emph{Data from other tables}: The query engine provides each table
adapter with the entire combined table shown to the user. The DOM
scraping adapter uses this for injecting annotations---values from other
tables are added to the original web page.

\emph{Currently selected record}: As the user clicks around the table
view, the query engine broadcasts the record currently selected by the
user to each table adapter. The DOM scraping adapter uses this
information to highlight the row in the page that corresponds to the
selected row in the table, which helps the user understand the mapping
between the table and the original UI.

Next, we present the three types of table adapters we have built in
Wildcard so far. These do not represent an exhaustive set of all
possible table adapters---in Section~\ref{sec:visions} we discuss other
types of adapters that would fit well into the general paradigm.
